---
author: 
title: About me
---

<img src="https://avatars0.githubusercontent.com/u/37358653?s=400&v=4" style="max-width:25%;min-width:40px;float:right;" alt="Churches avatar" />

## TL;DR
* Public sector statistician
* Former neuroscienctist

## Long version
### The past
I completed my PhD at the [Univeristy of Cambridge](https://www.autismresearchcentre.com/people_Churches) in 2011 and untill 2017, I was a neuroscientist. There is a good list of my academic publications [here](https://scholar.google.com/citations?user=nTXN-X8AAAAJ&hl=en). I quite enjoyed a lot about being an academic. I did some lectures where I lead the class in protest against the tyrany of frequentist statistics^[With appreciation and appologies to Bob Dylan]:

```{r echo = FALSE}
blogdown::shortcode("youtube", "2ze-_jA1X94")
```

And I was sometimes joined for lectures by a virtual teaching assistant.

```{r echo = FALSE}
blogdown::shortcode("youtube", "jDtSo6nijfI")
```

### The present
I am a statistician working for the Government of South Australia. I started this job in 2017. I provide analyses of administrative data for internal government communications and public facing data visualisations.

I also provide advice to government on the use of quantitative information including the construction of automated decision making systems. My work in this field led me to a [Churchill Fellowship](https://www.churchilltrust.com.au/fellows/detail/4329/Owen+Churches) in 2019 to look at ethical and just regulation of artificial intelligences.

I still collaborate with the [CAIN Laboratory](http://www.cain.science/) at UniSA. The beautiful knitted neurons on the landing page of this blog were made by the staff and students there. 



