---
title: South Australian Legislative Network
#author: Owen Churches
date: '2018-11-01'
slug: Legislative network
categories:
  - Legislation
tags:
  - legislation
  - networks
---
#Background
I joined the public service in 2017. Pretty soon after I started my boss and new colleagues suggested that the most important thing for me to read would be the legislation that I work under^[Which happens to be the [Children and Young People (Oversight and Advocacy Bodies) Act 2016](https://www.legislation.sa.gov.au/LZ/C/A/CHILDREN%20AND%20YOUNG%20PEOPLE%20(OVERSIGHT%20AND%20ADVOCACY%20BODIES)%20ACT%202016.aspx)]. 

The thing is, when reading *my* legislation, I realised that it referenced several other pieces of legislation. This meant that to fully understand *my* legislation I also had to read these referenced pieces of legislation. But guess what? All these pieces of legislation also referenced other pieces of legislation. I realised that this could go on for quite some time!

Also, I realised that this situation really suits a graph solution. In this sense:

* acts are nodes
* mentions of acts are links

My plan was:

1. Download all the legislation active in South Australia at the moment
2. Make a list of all the legislation
3. Look for mentions of other acts within each act and make an edge list
4. Make the graph
5. Explore the network metrics
6. Make an interactive plot of the whole thing because pictures of graphs are pretty and informative

#So, to work...
##Downloading the legislation
This was easier than I thought. South Australia is super progressive with its use of open data and there is a copy of all legislation [currently in use](https://data.sa.gov.au/data/dataset/legislative-database). Impressively, this is updated every two weeks^[And the previous updates are stored giving a snapshot of legislation active every two week. This raises the possibility of looking at the way legislation has evolved.]

##A list of legislation
The current legislation is in a folder called "A" (for "Acts"). 

```{r eval=FALSE}
library(here)
library(striprtf)
library(tidyverse)
library(tidygraph)
library(igraph)
library(ggraph)

acts <- list.files(path = here("A"), 
                        pattern = "rtf$", 
                        recursive = TRUE) %>%
  tibble() %>%
  rename("directory" = ".") %>%
  mutate(from = str_extract(directory, ".+?(?=/)"))
```

## Look for mentions
For some inexplicable reason the content of the acts is stored in rich text format. But `striprtf::read_rtf` works pretty well.
```{r eval=FALSE}
content <- ""
for (n in 1:nrow(acts)) {
  content[n] <- read_rtf(here("A", acts$directory[n])) %>%
    paste(collapse = " ")
}

content <- tibble(content)

acts <- bind_cols(acts, content)
```

## Make the graph

I tried and failed with a few differed regex approaches. Eventually I settled on this:

```{r eval=FALSE}
acts <- mutate(content = str_replace_all(content, "\\s+", " ")) %>% 
  mutate(to = str_extract_all(content, paste(as.character(from), collapse="|"))) %>%
  unnest(to) %>%
  select(from, to)

acts_graph <- as_tbl_graph(acts, directed = FALSE) %>% 
    mutate(community = as.factor(group_walktrap())) %>%
  mutate(centrality = centrality_degree())

ggraph(acts_graph, layout = 'graphopt', charge = .05) + 
    geom_edge_link(alpha = .8) + 
    geom_node_point(aes(colour = community, size = centrality), alpha = .3) + 
  guides(colour = FALSE) + 
    theme_graph() + 
  labs(title = str_wrap("Hairball of active legislation in South Australia November 2018", 50), caption = "Colour is the result of community detection and like the rest of the plot fulfils no explanatory purpose")
```

<img src = "/legislation_hairball.png" alt="" />

Ok, I should move on and never talk of that again.

## Explore the metrics
A decent place to start would be a histogram of the centrality of each act in the network.

```{r eval=FALSE}
acts_metrics <- acts_graph %>%
  activate(nodes) %>%
  as.tibble()

ggplot(data = acts_metrics, aes(x = centrality)) + 
  geom_histogram(binwidth = 2) + 
  labs(title = str_wrap("Histogram of centrality of acts in the network of legislation in South Australia, November 2018", 70))
```

<img src = "/legislation_histogram.png" alt="" />

So, there are way more acts with few connections than there are acts with many connection. Its not quite on the scale of a powerlaw though.

Something that would be interesting to check would be the relationship between the centrality of the acts and the year the act was assented to by the Governor.

```{r eval=FALSE}
acts_metrics <- acts_metrics %>%
  mutate(year = str_sub(name, -4, -1)) %>%
  mutate(year = as.numeric(year))

ggplot(data = acts_metrics, aes(x = year, y = centrality)) + 
  geom_point() +
  geom_label_repel(data = subset(acts_metrics, centrality > 30), 
                  aes(label = name, fill = community))
```
<img src = "/legislation_scatter.png" alt="" />

So the more recent acts are also the more interlinked acts. And from the change in density on the x axis, it seems that there are way more recent acts in the network. Also, yes, I snuck the community detection back in there for the highly connected acts and it seems that one of the communities deals with finance and the other with natural resources.

One other thing really stuck out is the line of acts that seem to be preserved in the current legislation that were all ascented to in 1936.
```{r eval = FALSE}
what_a_year <- acts_metrics %>%
  filter(year == 1936) %>% 
  select(name)

kable(what_a_year)
```
**name**  
Civil Liability Act 1936  
Debtors Act 1936  
Explosives Act 1936  
Land Tax Act 1936  
Landlord and Tennant Act 1936  
Law of Property Act 1936  
Limitation of Actions Act 1936  
Lottery and Gamy Act 1936  
Mercantile Act 1936  
Oaths Act 1936  
Redundant Officers Fund Act 1936  
Trustee Act 1936  
Wills Act 1936  

What happened in 1936 or what is it about these acts that has meant they have endured?

## Make an interactive plot

So, the last thing I wanted to do was make an interactive, zoomable, clickable network of the legislation. Serendipitously, I had just been introduced to the wonderful `networkD3` package via Twitter from some amazing network researchers at Adelaide University:

```{r echo = FALSE}
blogdown::shortcode("tweet", "1060706080374763520")
```

So [here](https://ofchurches.shinyapps.io/leg_app) it is. Its worth clicking through to open the `shinyapp` in its own window because its a big network. You can select the name of a piece of legislation to see its place in the network.